---
title: "Covid-19 vaccine study Final 2023"
author: "Dorothy Decker"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
guy = filter(FinalData, sex =="M")
gal = filter(FinalData, sex =="F")
gay = filter(FinalData, LGBTQ =="gay")
Druguser = filter(FinalData, DrugUser =="yes")
```


# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

## graphical males



```{r}
dd2 <- guy %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In this set of data, we are looking at the subgroup for males. We can see those who ended up infected were those who took the placebo instead of the vaccine. Over 60% of the males that took the placebo ended up infected. Looking at the results of the males who were fine, half of them took the placebo and half took the vaccine. There is a slight higher percentage of those who were okay that took the placebo. So far, it seems like the vaccine is working for males. 

## numerical males 

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
rowPerc(table1)
colPerc(table1)
```
In this set of numerical results, we can see the specific numbers of those who were affected. 69% of men that took the placebo got infected, while only 30% of men who took the vaccine got covid19. 

## inferential males

```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```

With all of this data, we can put our attention on the Fisher's Exact test. The p-value displayed above is extremely small at 5e-14. This is a really small number, the smaller the number the more it points that the vaccine may have an impact depending on gender. The 95 percent confidence interval is 1.8. 



The null hypothesis predicted that the deaths related to covid-19 are independent of sex. Since it is clear that sex does play a role in the death rate of individuals, the alternative hypothesis, that there is a correlation between covid-19 deaths and sex, is the more likely option.



```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate +1)*100
print(effectivenessguy)
```

```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options(digits=2)
```


Above is the effectiveness for the covid-19 vaccine for guys(males). If the effectiveness is above 50% that means that the vaccine is effective. We can see that it is as 69%. Which means it is effective for men.  The effectiveness confidence interval for the guys is 69.2, with a hi of 73.75 and a low of 64.26


## Females

## females graph 

```{r}
dd2 <- gal %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In this graph we can see that the females that took the placebo got infected with covid-19 more than those who took the vaccine. In the second category, we see those who were OK, and the bars are about the same.

## females numerical

```{r}
table1 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```

Looking at the numerical data, we can see that 60 of those who took the placebo got infected and only 40 who took the vaccine ended up getting covid-19. For the OK category, it was 50/50.

## females inferential

```{r}
chisq.test(table1)
chisqtestGC(table1)
galsfish <- fisher.test(table1)
galsfish
```

With all of this data, we can put our attention on the Fisher's Exact test. The p-value displayed above is small at 2e-04. This is a really small number, the smaller the number the more it points that the vaccine may have an impact depending on gender. The confidence interval is 1.2.


```{r}
effect <- galsfish$p.value/(galsfish$p.value +1)
effectivenessgal = galsfish$estimate/(galsfish$estimate +1)*100
print(effectivenessgal)
```

```{r}
low=(galsfish$conf.int[1]/(galsfish$conf.int[1] +1))*100
hi=(galsfish$conf.int[2]/(galsfish$conf.int[2] +1))*100
options(digits=2)
```

The effectiveness for the women are 59.9, with a hi of 65.07 and a low of 54.53 This is showing it is helping the females. 

## LGBTQ

## LGBTQ graph

```{r}
dd2 <- gay %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Looking at this set of graphs, it is very different. You can automatically see that there is a higher chance of those getting covid-19 when using the vaccine. While those who took the placebo are less likely to get infected. The OK side is still looking 50/50.


## Numerical LGBTQ


```{r}
table1 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

Looking at the numerical data, we can see that 15 of those who took the placebo were sick and 50 people were OK after taking the placebo. In the contrary 85 people who took the vaccine ended up getting infected, and 50 people were OK. This is beginning to show this vaccine may not be suitable due to sexuality.

## Inferential LGBTQ

```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```

Looking at the Fisher's Exact Test for count data, the P value is 1e-08. We can infer that this vaccine isn't as effective for this category of people within the sample. The confidence interval is 0.082.

```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgays = gaysfish$estimate/(gaysfish$estimate +1)*100
print(effectivenessgays)
```

```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))*100
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))*100
options(digits=2)
```

The effectiveness is very small, it is 15.4. The hi is 26.5 and a low of 7.60. This is not very effective for the category according to the data we have observed.


## Druggies

## Drug User Graph 


```{r}
dd2 <-  Druguser %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

According to the graphical data, we can see that those who took the vaccine ended up getting infected with covid-19 more than those who took the placebo. In the OK portion of the data, there is slightly more of drug users who ended up being OK vs those who took the vaccine. 

## Drug Users numerical data

```{r}
table1 <- xtabs(~infected + treatment, data= Druguser)
rowPerc(table1)
colPerc(table1)
```

Looking at the numerical data, we can further see that 88 of those who took the vaccine ended up infected, while 12 of them were those who took the placebo. In other words, majority of those who got infected were the ones who took the vaccine.

## Drug users inferential data


```{r}
chisq.test(table1)
chisqtestGC(table1)
drugfish <- fisher.test(table1)
drugfish
```

Once again, if we pay attention to the Fisher's Exact Test for count data, we can see its 2e-10.  The 95 percent confidence interval is 0.048. This shows that the vaccine does not work for the group of Drug Users within this data set. 


```{r}
effect <- drugfish$p.value/(drugfish$p.value +1)
effectivenessdrug = drugfish$estimate/(drugfish$estimate +1)*100
print(effectivenessdrug)
```

```{r}
low=(drugfish$conf.int[1]/(drugfish$conf.int[1] +1))*100
hi=(drugfish$conf.int[2]/(drugfish$conf.int[2] +1))*100
options(digits=2)
```

The odds ratio is 11, which the exact effectivness for drug users is 11.2. The hi is 21.82 and the low is 4.5.


# Overall Results and Conclusions